---
title: "Biostat M280 Homework 4"
subtitle: Due Mar 16 @ 11:59PM
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(sparklyr)
library(dplyr)
library(ggplot2)
Sys.setenv(SPARK_HOME="/usr/lib/spark")
config <- spark_config()
sc <- spark_connect(master = "yarn-client", config = config)
#spark_disconnect_all()
```
* Prepare the database. 
```{r}
#Access Hive tables
flights_tbl <- tbl(sc, 'flights')
#flights_tbl %>% print(width = Inf)
airlines_tbl <- tbl(sc, 'airlines')
airports_tbl <- tbl(sc, 'airports')
#airports_tbl %>% print(width = Inf)
flight2000 <- flights_tbl %>% 
  filter(year == 2000)
#flight2000 %>% print(width = Inf)
LAX <- flights_tbl %>% 
  filter(origin == "LAX" || dest == "LAX") 
LAX_traf <- LAX %>% 
  group_by(year, month, dayofmonth, dayofweek) %>%
  summarise(n = n()) %>%
  collect()
LAX2000_dest <- flight2000 %>%
  filter(origin == "LAX") %>%
  group_by(dest) %>%
  summarise(n = n()) %>%
  arrange(desc(n)) %>%
  head(10)
#LAX %>% print(width = Inf)

#Find the most departure delay dest from LAX
LAX_delay <- flight2000 %>%
  filter(origin == "LAX") %>%
  filter(depdelay >= 0) %>%
  group_by(dest) %>%
  summarise(avgdelay = mean(depdelay, na.rm = T)) %>%
  arrange(desc(avgdelay)) %>%
  head(10)
LAX_delay30 <- flight2000 %>% 
  filter(origin == "LAX") %>%
  filter(depdelay >= 30) %>%
  group_by(dest) %>%
  summarise(n = n()) %>%
  arrange(desc(n)) %>%
  head(10)
LAX_delay30 %>% print(width = Inf)
```

Our Apache Yarn cluster hosts the [flights](http://stat-computing.org/dataexpo/2009/the-data.html) data representing 123 million flights over 22 years. Read the [lecture notes](http://hua-zhou.github.io/teaching/biostatm280-2018winter/slides/12-sparklyr/sparklyr-flights.html) on how to access the Yarn cluster. Connect to the database using `sparklyr` and answer following questions. You can base your answers on a specific year or the whole data set.
#### My project was based on the data in year 2000.
1. Map the top 10 busiest airports. Size of dots should reflect the number of flights through that destination.  
Hint: You may find this tutorial on [Making Maps in R](http://eriqande.github.io/rep-res-web/lectures/making-maps-with-R.html) helpful.
* The data was based on the year *2000*. The busy airports were based on the number of flights through that *destination*.
```{r}
# map pakeages 
library(ggplot2)
library(ggmap)
library(maps)
library(mapdata)

#find top 10 busy airport in year 2000
busy <- flight2000 %>%
  group_by(dest) %>% 
  summarise(faa = dest, n = n()) %>% 
  arrange(desc(n)) %>%
  head(10) %>% 
  collect()
# innerjoin by airport name (faa)
busy_ap <- inner_join(busy, airports_tbl, by = "faa", 
                      copy = T) %>%
            arrange(desc(n))
busy_ap$lat <- as.numeric(busy_ap$lat)
busy_ap$lon <- as.numeric(busy_ap$lon)

usa <- map_data("usa")
gg1 <- ggplot() + 
  geom_polygon(data = usa, aes(x = long, y = lat, group = group),
               fill = "navajowhite", color = "navajowhite4") + 
    coord_fixed(1.3) + 
  geom_point(data = busy_ap, aes(x = lon, y = lat, size = n )) + 
  geom_text(data = busy_ap, aes(lon, lat, label=dest,
                                colour = factor(dest), fontface = "bold"))
gg1
```

2. Map the top 10 busiest direct routes. Size of lines should reflect the number of flights through that route.
* The data based on the year 2000. 
```{r}
busyroutes <- flight2000 %>%
  group_by(origin,dest) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n)) %>%
  head(10) %>% 
  collect()
toproutes <- busyroutes %>% 
  left_join(airports_tbl, by = c("origin" = "faa"), copy = T) %>% 
  rename(lon_from = lon, lat_from = lat) %>% 
  left_join(airports_tbl, by = c("dest" = "faa"), copy = T) %>% 
  rename(lon_to = lon, lat_to = lat) %>%
  select(origin, lon_from, lat_from, dest, lon_to, lat_to, n)

# Convert the lat and lon from character to double
toproutes$lon_from <- as.numeric(toproutes$lon_from)
toproutes$lon_to <- as.numeric(toproutes$lon_to)
toproutes$lat_from <- as.numeric(toproutes$lat_from)
toproutes$lat_to <- as.numeric(toproutes$lat_to)
toproutes$scale_n <- 0.1 + 1.9 * (
  toproutes$n - min(toproutes$n))  / (max(toproutes$n)- min(toproutes$n))

airport <- unique(c(toproutes$origin, toproutes$dest)) 
topairports <- as.data.frame(airport)
topairports <- topairports %>%
  left_join(airports_tbl, by = c("airport" = "faa"), copy = T) %>% 
  select(airport, lon, lat)
topairports$lon <- as.numeric(topairports$lon)
topairports$lat <- as.numeric(topairports$lat)
topairports

```

Plot

```{r}
usa <- map_data("usa")
gg2 <- ggplot() + 
  geom_polygon(data = usa, aes(x = long, y = lat, group = group),
               fill = "azure2", color = "azure4") + 
    geom_curve(data = toproutes, aes(x = lon_from, xend = lon_to, y = lat_from, yend = lat_to, colour = factor(n)), curvature = .5, arrow = arrow(length = unit(0.3, "cm")), size = toproutes$scale_n) +
    coord_fixed(1.3) + 
  geom_point(data = topairports, aes(x = lon, y = lat))  
  #geom_text(data = topairports, aes(lon, lat, label=airport, fontface = "bold"), colour = "red")

gg2
```


3. LAX:
  

    (a). Reproduce above plot. Visualize and explain some prominent features you observe. For example, what happened at points 1-5?
```{r}
LAX_traf$date <- as.Date(paste(LAX_traf$year, LAX_traf$month, LAX_traf$dayofmonth, sep='-'))

break.vec <- seq(from=as.Date("1998-01-01"), to=as.Date("2008-01-01"), by="year")
gg3 <- ggplot() +
  geom_line(data = LAX_traf, aes(x = date, y = n)) + 
  scale_x_date(breaks = break.vec,date_labels = "%Y", 
               limits = as.Date(c('1998-1-1', '2009-1-1'))) +
  coord_fixed(ratio=1.2) +
  ggtitle("LAX air traffic") +  
  geom_label(label = c("5","1","3","2","4"), 
             aes(x = as.Date(c('2001-1-1', '2001-9-11', 
                               '2004-7-1', '2004-11-25', '2008-1-1')), 
             y = c(1200, 1070, 960, 930, 1200)))
gg3
```
* Explaination: The label #1 showed sharply decrease, I believed the reason is the 911 make people don't want to travel a lot. The label #2 and #3 showed the in Noverber (or December) and July there is holidays make this parttern. Since during Thanksgiving or close to Chrismas, families member usually want to stay at home and travel less. Also, independence day in July, people won't travel for business a lot. The label #4 showed sharply decrease, the possible reason for this is because of economic depression in 2008. The label #5 showed increase is because of Los Angeles finished city constructions, so people travel to LA more. 

    (b). Visualize and explain seasonal effects.
    * Answer based on the flight in 2000
```{r}
LAX2000 <- LAX_traf %>% 
  filter(year == 2000)

LAX2000$season <- (LAX2000$month%/%4 + 1)
#table to show seasonal difference
LAX2000 %>% 
  group_by(season) %>%
  summarise(traffic_n = sum(n))
#plot to show seasonal difference
gg4 <- ggplot() +
  geom_line(data = LAX2000, aes(x = date, y = n)) + 
  scale_x_date(date_breaks = "1 month", 
               date_labels = "%m", limits = as.Date(c('2000-1-1', 
                                                      '2000-12-31'))) +
  coord_fixed(ratio=0.2) +
   xlab("Date") + ylab("Number of flights") + 
  ggtitle("Seasonal LAX air traffic in 2000")
gg4
```
* The table and plot showed the seasonal effect on the flight traffic in LAX (in 2000). Winter quater had lowest traffic and summer and fall quater had busies traffic. The possible reason for this is because students have their summer break so they have their vacation during summer. In addition, people usually tend to stay at home during winter due to holiday and bad weather. 

    (c). Visualize and explain weekly effects.
    * Answer based on the flight in 2000
```{r}
week <- LAX2000 %>%
  group_by(dayofweek) %>%
  summarise(traffic_n = sum(n))
ggplot() +
  geom_bar(data = week, aes(x = dayofweek, y = traffic_n), stat = "identity")
```
  * The Sat and Sun had less traffic. From Monday to Friday, there were high demands for flights for business travel.  
  
    (d). Map top 10 destinations from LAX. Size of dots should reflect the number of flights from LAX to that destination.
    * Answer `based on the flight in 2000`. 
```{r}
LAX2000_dest1 <- inner_join(LAX2000_dest, airports_tbl,  
                        by = c("dest" = "faa"), copy = T) %>%
              ungroup() %>%
             select(dest, n, lat, lon) %>%
              collect()
```
```{r}
LAX2000_dest1$lat <- as.numeric(LAX2000_dest1$lat)
LAX2000_dest1$lon <- as.numeric(LAX2000_dest1$lon)
usa <- map_data("usa")
gg5 <- ggplot() + 
  geom_polygon(data = usa, aes(x = long, y = lat, group = group),
               fill = "seashell", color = "seashell3") + 
    coord_fixed(1.3) + 
  geom_point(data = LAX2000_dest1, aes(x = lon, y = lat, 
                                       size = n, color = dest))
  
gg5
```
  
4. Build a predictive model for the arrival delay (`arrdelay`) of flights flying from LAX. Use the same filtering criteria as in the [lecture notes](http://hua-zhou.github.io/teaching/biostatm280-2018winter/slides/12-sparklyr/sparklyr-flights.html) to construct training and validation sets. You are allowed to use a maximum of 5 predictors. The prediction performance of your model on the validation data set will be an important factor for grading this question.
### Compute the linear model.
```{r}
system.time(
model_data <- flights_tbl %>%
    filter(origin == "LAX") %>%
    filter(!is.na(arrdelay) & !is.na(depdelay) & !is.na(distance)) %>%
    filter(depdelay > 15 & depdelay < 240) %>%
    filter(arrdelay > -60 & arrdelay < 360) %>%
    filter(year >= 2003 & year <= 2007) %>%
    left_join(airlines_tbl, by = c("uniquecarrier" = "code")) %>%
    mutate(hour = as.integer(crsdeptime / 100)) %>%
    select(year, month, dayofweek, arrdelay, depdelay, 
           crsdeptime, distance, uniquecarrier, description, hour)
)

# Partition the data into training and validation sets
model_partition <- model_data %>% 
  sdf_partition(train = 0.8, valid = 0.2, seed = 5555)
system.time(
  ml1 <- model_partition$train %>%
    ml_linear_regression(arrdelay ~ distance + depdelay + dayofweek + hour)
)
# Summarize the linear model
summary(ml1)
#Find MSE
result_valid <- sdf_predict(ml1, model_partition$valid) %>%
  mutate(residual = arrdelay - prediction) %>%
  summarize(mse = mean(residual^2, na.rm = T))
result_valid
```
* The R-Square: *0.8834*.
* The MSE: Model: 13.44^2 = *180.6336*, Prediction: *181.9295*.
* The R-Square is 0.8834, which close to 1. In addition, the MSE of model and valid prediction were very close. Therefore, the model (arrdelay ~ distance + depdelay + dayofweek + hour) is reliable.

### Assess model performance
```{r}
# Calculate average gains by predicted decile
system.time(
  model_deciles <- lapply(model_partition, function(x) {
    sdf_predict(ml1, x) %>%
      mutate(decile = ntile(desc(prediction), 10)) %>%
      group_by(decile) %>%
      summarize(arrdelay = mean(arrdelay)) %>%
      select(decile, arrdelay) %>%
      collect()
  })
)
model_deciles
# Create a summary dataset for plotting
deciles <- rbind(
  data.frame(data = 'train', model_deciles$train),
  data.frame(data = 'valid', model_deciles$valid),
  make.row.names = FALSE
)
deciles

# Plot average arrdelay by predicted decile
deciles %>%
  ggplot(aes(factor(decile), arrdelay, fill = data)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  labs(title = 'Average arrdelay by predicted decile', 
       x = 'Decile', y = 'Minutes')
```

* Based on the Plot above, the difference be

### Visualize predictions
```{r}
# Select data from an out of time sample
data_2000 <- flights_tbl %>%
  filter(origin == "LAX") %>%
  filter(!is.na(arrdelay) & !is.na(depdelay) & !is.na(distance)) %>%
  filter(depdelay > 15 & depdelay < 240) %>%
  filter(arrdelay > -60 & arrdelay < 360) %>%
  filter(year == 2000) %>%
  left_join(airlines_tbl, by = c("uniquecarrier" = "code")) %>%
  mutate(hour = as.integer(crsdeptime / 100)) %>%
  select(year, month, dayofweek, arrdelay, depdelay, crsdeptime, 
         distance, uniquecarrier, description, hour)
data_2000
```
```{r}
# Summarize data by carrier
carrier <- sdf_predict(ml1, data_2000)  %>%
   group_by(description) %>%
   summarize(arrdelay = mean(arrdelay, na.rm = T), 
             prediction = mean(prediction, na.rm = T), freq = n()) %>%
   filter(freq > 100) %>%
  collect()
hour <- sdf_predict(ml1, data_2000)  %>%
   group_by(hour) %>%
   summarize(arrdelay = mean(arrdelay, na.rm = T), 
             prediction = mean(prediction, na.rm = T), freq = n()) %>%
   filter(freq > 300) %>%
  collect()
```

```{r}
# Plot actual and predicted arrival delay by airline carrier
ggplot(carrier, aes(arrdelay, prediction)) + 
  geom_point(alpha = 0.75, color = 'red', shape = 3) +
  geom_abline(intercept = 0, slope = 1, alpha = 0.15, color = 'blue') +
  geom_text(aes(label = substr(description, 1, 20)), 
            size = 3, alpha = 0.75, vjust = -1) +
  labs(title='Average Arrival Delay Forecast', x = 'Actual', y = 'Predicted')

# Plot actual and predicted arrival delay by Hour
ggplot(hour, aes(arrdelay, prediction)) + 
  geom_point(alpha = 0.75, color = 'red', shape = 3) +
  geom_abline(intercept = 0, slope = 1, alpha = 0.15, color = 'blue') +
  geom_text(aes(label = substr(hour, 1, 20)), 
            size = 3, alpha = 0.75, vjust = -1) +
  labs(title='Average Arrival Delay Forecast', x = 'Actual', y = 'Predicted')
```
```{r}
# # Prediction results for model 2
# carrier2 <- sdf_predict(ml2, data_2000)  %>%
#    group_by(description) %>%
#    summarize(arrdelay = mean(arrdelay, na.rm = T), prediction = mean(prediction, na.rm = T), freq = n()) %>%
#    filter(freq > 100) %>%
#   collect()
# ```
# 
# ```{r}
# # Plot actual gains and predicted gains by airline carrier
# ggplot(carrier2, aes(arrdelay, prediction)) + 
#   geom_point(alpha = 0.75, color = 'red', shape = 3) +
#   geom_abline(intercept = 0, slope = 1, alpha = 0.15, color = 'blue') +
#   geom_text(aes(label = substr(description, 1, 20)), size = 3, alpha = 0.75, vjust = -1) +
#   labs(title='Average Arrival Delay Forecast', x = 'Actual', y = 'Predicted')
```

    
5. Visualize and explain any other information you want to explore.
* I choose to find the top 10 destination that most frequently delay more than 30 min.
```{r}
LAX_delay <- inner_join(LAX_delay30, airports_tbl,
                          by = c("dest" = "faa"), copy = T) %>%
  ungroup() %>%
  select(dest, n, lat, lon) %>%
  collect()
```
```{r}

LAX_delay$lat <- as.numeric(LAX_delay$lat)
LAX_delay$lon <- as.numeric(LAX_delay$lon)
usa <- map_data("usa")
gg6 <- ggplot() + 
  geom_polygon(data = usa, aes(x = long, y = lat, group = group),
               fill = "honeydew", color = "honeydew4") + 
    coord_fixed(1.3) + 
  geom_point(data = LAX_delay, aes(x = lon, y = lat, color = n, size = 3)) +
   scale_color_gradientn(colours=rainbow(3))
gg6
```


### Disconnect 
```{r}
spark_disconnect_all()
```
  
  