---
title: "Biostat M280 Homework 2"
subtitle: Due Feb 16 @ 11:59PM
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Q1

Read [Chapter 7](http://r4ds.had.co.nz/exploratory-data-analysis.html) (Exploratory Data Analysis) of _R for Data Science_ and do exercises 7.3.4, 7.4.1, 7.5.1.1, 7.5.2.1, and 7.5.3.1.

## Q2 (optional)

Read [Chapter 23](http://r4ds.had.co.nz/model-basics.html) (Model Basics) and [Chapter 24](http://r4ds.had.co.nz/model-building.html) (Model Building) of _R for Data Science_ and do exercises 24.2.3 and 24.3.5.

## Q3

Redo HW1 Q2 using tidyverse.
* Read in data
```{r}
library("tidyverse")
geno_data1 <- read_tsv("/home/m280-data/hw1/merge-geno.bim", col_names = F) 
geno_data1 <- as_data_frame(geno_data1)
colnames(geno_data1) <- c("Chromosome","SNP_ID","Genetic_Distance","bp",
                       "Allele 1","Allele 2")
geno_data2 <- read_delim("/home/m280-data/hw1/merge-geno.fam"," ", col_names = F)
geno_data2 <- as_data_frame(geno_data2)
colnames(geno_data2) <- c("Family","Person","Father","Mother",
                       "Sex","Affection")
```
1. How many persons are in the data set (statisticians call this n)? How many SNPs are in the data set (statisticians call this p)?

```{r}
geno_data1 %>% 
  count()
geno_data2 %>% 
  count()
```
* Answer: n = $959$ ; p = $8348674$. 

2. Which chromosomes does this data set contain? How many SNPs are in each chromosome?

```{r}
geno_data1 %>% 
  group_by(Chromosome) %>% 
  summarise(n = n()) 
```
* Answer:

| $Chromesome$ | $SNPs$ |
|--------------|--------|
|     1        |1309299 |
|     3        |1215399 |
|     5        |1090185 |
|     7        | 980944 |
|     9        | 732013 |
|     11       | 815860 |
|     13       | 602809 |
|     15       | 491208 |
|     17       | 477990 |
|     19       | 393615 |
|     21       | 239352 |



3. MAP4 (microtubule-associated protein 4) is a gene on chromosome 3 spanning positions 47,892,180 bp – 48,130,769 bp. How many SNPs are located within MAP4 gene?
```{r}
filter(geno_data1, Chromosome == 3, bp >= 47892180, bp <= 48130769 ) %>% 
  summarise(n = n()) 
```
* Answer: There are 894 SNPs located within MAP4 gene.

4. Reformat:
* 

## Q4 (optional)

Redo HW1 Q3 on Hoffman2, except now we want to submit each `runSum.R` job to a different node in the cluster.
